---
title: "CodeBook"
author: "Roberto B Paixao"
date: "05/07/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The steps I did in the file run_analysis.R

- First, activated the package: 
```{r}
library(dplyr)
```

- Download the dataset and unzip it:
```{r}
if (!file.exists(filename)){
    fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
    download.file(fileURL, filename, method="curl")
}  
unzip(filename)
```

- Create the data frames:
The data frames were created separately using read.tbale function.
```{r}
features <- read.table("UCI HAR Dataset/features.txt", col.names = c("n","functions"))
activities <- read.table("UCI HAR Dataset/activity_labels.txt", col.names = c("code", "activity"))
subject_test <- read.table("UCI HAR Dataset/test/subject_test.txt", col.names = "subject")
x_test <- read.table("UCI HAR Dataset/test/X_test.txt", col.names = features$functions)
y_test <- read.table("UCI HAR Dataset/test/y_test.txt", col.names = "code")
subject_train <- read.table("UCI HAR Dataset/train/subject_train.txt", col.names = "subject")
x_train <- read.table("UCI HAR Dataset/train/X_train.txt", col.names = features$functions)
y_train <- read.table("UCI HAR Dataset/train/y_train.txt", col.names = "code")
```

- 1. To merge
Using rbind function the X, Y, and Subjetc data sets were merged.
```{r}
X <- rbind(x_train, x_test)
Y <- rbind(y_train, y_test)
Subject <- rbind(subject_train, subject_test)
Merged_Data <- cbind(Subject, Y, X)
```

# 2. Extractions
The function Select was used to filter only the variables with "mean" and "std".
```{r}
TidyData <- Merged_Data %>% 
    select(subject, code, contains("mean"), contains("std"))
```

# 3. Activities names
The names were assigned to the activities on variable code.
```{r}
TidyData$code <- activities[TidyData$code, 2]
```

# 4. Labels
The labels were organized using the gsub function to be more undestandable.
```{r}
names(TidyData)[2] = "activity"
names(TidyData)<-gsub("Acc", "Accelerometer", names(TidyData))
names(TidyData)<-gsub("Gyro", "Gyroscope", names(TidyData))
names(TidyData)<-gsub("BodyBody", "Body", names(TidyData))
names(TidyData)<-gsub("Mag", "Magnitude", names(TidyData))
names(TidyData)<-gsub("^t", "Time", names(TidyData))
names(TidyData)<-gsub("^f", "Frequency", names(TidyData))
names(TidyData)<-gsub("tBody", "TimeBody", names(TidyData))
names(TidyData)<-gsub("-mean()", "Mean", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("-std()", "STD", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("-freq()", "Frequency", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("angle", "Angle", names(TidyData))
names(TidyData)<-gsub("gravity", "Gravity", names(TidyData))
```

# 5. Second data set
The data were grouped by Subject and activity using the function group_by and summarise_all with means.
The final txt file was generated using the function write.table.
```{r}
FinalData <- TidyData %>%
    group_by(subject, activity) %>%
    summarise_all(funs(mean))
write.table(FinalData, "FinalData.txt", row.name=FALSE)
```

